Batch, Off-policy and Model-free Apprenticeship Learning

### 2011-06-21 main.tex
Version soumise à EWRL

\usepackage{amssymb}
\usepackage{natbib}

@@

\title{Apprentissage par imitation dans un cadre \emph{batch}, \emph{off-policy} et sans modèle}
%\author{Robert}
%   1. Supelec-Metz Campus, IMS Reasearch group, France. \texttt{prenom.nom@supelec.fr}}
\author{Edouard Klein\inst{1}\inst{3} \and Matthieu Geist\inst{1} \and Olivier Pietquin\inst{1}\inst{2}}
 \institute{
 1. Supélec-Metz Campus, IMS Research group, France, \texttt{prenom.nom@supelec.fr}\\
 2. UMI 2958 CNRS - GeorgiaTech, France\\
 3. Equipe ABC, LORIA-CNRS, France.
 }
\date{\today}

@@

\section{Background}
\label{sec:back}
A sequential decision problem is often framed as a Markov Decision
Process (MDP)~\citep{puterman1994markov}, a tuple
$\{S,A,P,R,\gamma\}$ with $S$ being the state space, $A$ the action
space, $P\in\mathcal{P}(S)^{S\times A}$ the set of Markovian
transition probabilities, $R\in\mathbb{R}^S$ the reward function
(assumed to be absolutely bounded by 1) and $\gamma\in[0,1[$ a
discounting factor. A policy $\pi\in A^S$ maps states to actions. The
quality of a policy is quantified by the associated value function
$V^\pi$, which associates to each state the expected and discounted
cumulative reward:
\begin{equation}
  V^\pi(s) = E[\sum_{t=0}^\infty \gamma^t R(s_t)|s_0=s, \pi]
\end{equation}
Dynamic programming aims at finding the optimal policy $\pi^*$, that
is one of the policies associated with the optimal value function,
$V^* = \argmax_\pi V^\pi$, which maximizes the value for each state.
If the model (that is transition probabilities and the reward
function) is unknown, learning the optimal control through
interactions is addressed by RL.

@@

We assume that the true reward function belongs to some hypothesis
space $\mathcal{H}_\phi = \{\theta^T \phi(s),
\theta\in\mathbb{R}^p\}$, of which we assume the basis functions to
be bounded by 1: $|\phi_i(s)|\leq 1, \forall s\in S, 1\leq i \leq
p$. Therefore, there exists a parameter vector $\theta^*$ such that:
\begin{equation}
  R^*(s) = (\theta^*)^T \phi(s)
\end{equation}
In order to ensure that rewards are bounded, we impose that
$\|\theta\|_2\leq 1$. For any reward function belonging to
$\mathcal{H}_\phi$ and for any policy $\pi$, the related value
function $V^\pi(s)$ can be expressed as follows:
\begin{align}
  V^\pi(s) &= E[\sum_{t=0}^\infty \gamma^t \theta^T \phi(s_t)|s_0=s, \pi]
  \\
  &= \theta^T  E[\sum_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{align}
Therefore, the value function is also linearly parameterized, with
the same weights and with basis functions being grouped into the
so-called \emph{feature expectation} $\mu^\pi$:
\begin{equation}
  \mu^\pi(s) = E[\sum_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{equation}
Recall that the problem is to find a policy whose performance is
close to that of the expert's for the starting state $s_0$, on the
unknown reward function $R^*$. In order to achieve this goal, it is
proposed in~\citep{abbeel2004apprenticeship} to find a policy $\tilde{\pi}$ such
that $\|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2\leq \epsilon$
for some (small) $\epsilon>0$. Actually, this ensures that the value
of the expert's policy and the value of the estimated policy (for
the starting state $s_0$) are close for \emph{any} reward function
of $\mathcal{H}_\phi$:
\begin{align}
  |V^{\pi_E}(s_0) - V^{\tilde{\pi}}(s_0)| &=
  |\theta^T(\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0))|
  \label{eqn:vs0}
  \\
  &\leq \|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2
\end{align}
This equation uses the Cauchy-Schwarz inequality and the
assumption that $\|\theta\|_2\leq 1$. Therefore, the approach
described here does not ensure to retrieve the true reward function
$R^*$, but to act as well as the expert under this reward function
(and actually under any reward function).

@@

As suggested in~\citep{abbeel2004apprenticeship}, solving the MDP can be done
approximately by using any appropriate reinforcement learning
algorithm. In this paper, we use the Least-Squares Policy Iteration
(LSPI) algorithm~\citep{lagoudakis2003least}. There remains to estimate
feature expectations. In~\citep{abbeel2004apprenticeship}, $\mu^{\pi_E}(s_0)$ is
estimated using a Monte Carlo approach over $m$ trajectories:
\begin{equation}
  \hat{\mu}_E(s_0) = \frac{1}{m} \sum_{h=1}^m \sum_{t=0}^\infty
  \gamma^t \phi(s_t^{(h)})
\end{equation}
This approach does not hold if only transitions of the expert are
available, or if trajectories are too long (in this case, it is
still possible to truncate them). For intermediate policies, it is
also suggested to estimate associated feature expectations using a
Monte Carlo approach (if they cannot be computed exactly). This is
more constraining than for the expert, as this assumes that a
simulator of the system is available. In order to address these
problems, we introduce a temporal-difference-based algorithm to
estimate these feature expectations.

\section{LSTD-$\mu$}
\label{sec:lstdmu}
Let us write the definition of the $i^\text{th}$ component of a
feature expectation $\mu^\pi(s)$ for some policy $\pi$:
\begin{equation}
  \mu_i^\pi(s) = E[\sum_{t=0}^\infty \gamma^t \phi_i(s_t)|s_0=s,\pi]
  \label{eqn:phi}
\end{equation}
This is exactly the definition of the value function of the policy
$\pi$ for the MDP considered with the $i^\text{th}$ basis function
$\phi_i(s)$ as the reward function. There exist many algorithms to
estimate a value function, any of them can be used to estimate
$\mu_i^\pi$. Based on this remark, we propose to use specifically
the least-squares temporal difference (LSTD)
algorithm~\citep{bradtke1996linear} to estimate each component of the
feature expectation (as each of these components can be understood
as a value function related to a specific and known reward
function).

Assume that a set of transitions $\{(s_t,r_t,s_{t+1})_{1\leq t \leq
n}\}$ sampled according to the policy $\pi$ is available. We assume
that value functions are searched for in some  hypothesis space
$\mathcal{H}_\psi = \{ \hat{V}_\xi(s) = \sum\limits_{i=1}^q \xi_i \psi_i(s)
= \xi^T \psi(s), \xi\in\mathbb{R}^q\}$. As reward and value
functions are possibly quite different, another hypothesis space is
considered for value function estimation. But if $\mathcal{H}_\phi$
is rich enough, one can still consider
$\mathcal{H}_\psi=\mathcal{H}_\phi$. Therefore, we are looking for
an approximation of the following form:
\begin{equation}
  \hat{\mu}_i^\pi(s) = (\xi_i^*)^T \psi(s)
  \label{eqn:psi}
\end{equation}
The parameter vector $\xi_i^*$ is here the LSTD estimate:
\begin{equation}
  \xi_i^* = \left(\sum_{t=1}^n
  \psi(s_t)(\psi(s_t)-\gamma\psi(s'_{t}))^T\right)^{-1}
  \sum_{t=1}^n \psi(s_t) \phi_i(s_t)
\end{equation}
For apprenticeship learning, we are interested more particularly in
$\hat{\mu}^\pi(s_0)$. Let $\Psi = (\psi_i(s_t))_{t,i}$ be the
$n\times q$ matrix of values predictors, $\Delta\Psi = (\psi_i(s_t)
- \gamma\psi_i(s'_t))_{t,i}$ be the related $n\times q$ matrix and
$\Phi = (\phi_i(s_t))_{t,i}$ the $n\times p$ matrix of reward
predictors. It can be easily checked that $\hat{\mu}^\pi(s_0)$
satisfies:
\begin{equation}
  (\hat{\mu}^\pi(s_0))^T = \psi(s_0)^T (\Psi^T \Delta\Psi)^{-1} \Psi^T \Phi
\end{equation}
This provides an efficient way to estimate the feature expectation
of the expert in $s_0$.

There remains to compute the feature expectations of intermediate
policies, which should be done in an off-policy manner (that is
without explicitly sampling trajectories according to the policy of
interest). To do so, still interpreting each component of the
feature expectation as a value function, we introduce a state-action
feature expectation defined as follows (much as the classic
$Q$-function extends the value function):
\begin{equation}
  \mu^\pi(s,a) = E[\sum_{t=0}^\infty \gamma^t
  \phi(s_t)|s_0=s,a_0=a,\pi]
\end{equation}
Compared to the classic feature expectation, this definition adds a
degree of freedom on the first action to be chosen before following
the policy $\pi$. With a slightly different definition of the
related hypothesis space, each component of this feature expectation
can still be estimated using the LSTD algorithm (namely using the
LSTD-Q algorithm~\citep{lagoudakis2003least}). The clear advantage of
introducing the state-action feature expectation is that this
additional degree of freedom allows off-policy learning.
Extending LSTD-$\mu$ to state-action LSTD-$\mu$ is done in the same
manner as LSTD is extended to LSTD-Q~\citep{lagoudakis2003least},
technical details are not provided here for the clarity of
exposition.

@@

We want to compare the efficiency of both versions of the algorithm with respect to the number of samples available from the expert, as these samples usually are the bottleneck. Indeed as they are quite costly (in both means and human time) to generate they are often not in abundance hence the critical need for an algorithm to be expert-sample efficient. Having a simulator at one's disposal can also be difficult. For the simple problem we use this is however not an issue. The discussion about the choice of the performance metric has its own dedicated subsection (subsection~\ref{ssec:quality}). We use here the $||\mu^{\pi^{(j)}}(s_0) - \mu^{\pi_E}(s_0)||_2$ error term. Fig.~\ref{fig:E} shows, for some numbers of samples from the expert, the value of $||\mu^{\pi^{(j)}}(s_0)-\mu^{\pi_E}(s_0)||_2$ after the algorithm converged or attained the maximum number of iterations (we used 40). The best policy is found by LSTD variant after one iteration only\footnote{Precise reasons why it happens are not clear now, but certainly have something to do with the fact that all the estimations are made along the same distribution of samples.} whereas in the MC variant, convergence happens after at least ten iterations. The best policy is not always the last, and although it has always been observed so far with the LSTD variant, there is absolutely no way to tell whether this is a feature. The score of the best policy (not the last) according to the $||\mu^{\pi^{(j)}}(s_0) - \mu^{\pi_E}(s_0)||_2$ error term is plotted here. We can see that although the LSTD variant is not as good as the MC variant when very few samples are available, both algorithms quickly converge to almost the same value ; our version converges to a slightly lower error value. The fact that our variant can work in a batch, off-policy and model-free setting should make it suitable to a range of tasks inaccessible to the MC variant, the requirement of a simulator being often constraining.
\subsubsection{Inverted Pendulum}
\begin{figure}
\centering
\resizebox{\columnwidth}{!}{\input{../Code/InvertedPendulum/threshold}}
\caption{Number of balancing steps for the policies found by both variants with respect to the number of samples from a sub-optimal policy given to LSPI. Note the absence of middle ground in the quality.}
\label{fig:threshold}
\end{figure}
\begin{figure}
\centering
\resizebox{\columnwidth}{!}{\input{../Code/InvertedPendulum/threshold_EB}}
\caption{Same plot as Fig.~\ref{fig:threshold}, with mean and standard deviation over 100 runs. The big standard deviation stems from the absence of middle ground in the quality of the policies. The increasing mean w.r.t. the abscissa means that the proportion of good policies increases with the number of samples from a sub-optimal policy given to LSPI.}
\label{fig:threshold_EB}
\end{figure}
Another classic toy problem is the \emph{inverted pendulum}, used for example in \citep{lagoudakis2003least} from where we drew the exact configuration (mass of the pendulum, length of the arm, definition of the feature space, etc.). In this setting, the machine must imitate an expert who maintains the unstable equilibrium of an inverted pendulum allowed to move along only one axis. A balancing step is considered successful (reward $0$) when the angle is less than $\pi\over 2$. It is considered a failure (reward $-1$) when the pendulum has fallen (i.e. the angle exceeds $\pi \over 2$). A run is stopped after a failure or after 3000 successful balancing steps.\\

@@

Fig.~\ref{fig:threshold_EB} shows that when given samples from a sub-optimal policy, the LSTD variant can sort the problem out, statistically almost always with 3000 sub-optimal samples, and sometimes with as low as 200 sub-optimal samples. Such a setting is still batch, off-policy and model-free. When given enough sub-optimal samples, both variants are successful with only one trajectory from the expert (i.e. 100 samples, this is what is plotted here). Giving more trajectories from the expert does not improve the success rate.
\subsection{Discussion about the quality criterion}
\label{ssec:quality}
\begin{figure}
\centering
\input{../Code/GridWorld/criteria_mc}
\caption{Different criteria with respect to the number of iterations for a run of the MC variant.}
\label{fig:A}
\end{figure}
\begin{figure}
\centering
\input{../Code/GridWorld/criteria_lstd_EB}
\caption{Different criteria with respect to the number of samples from the expert, for several runs of the LSTD variant. We can see that the algorithm is blind, as all it has access to is always zero. The true error values, however, smoothly converge to something small. Knowing how many expert samples are actually needed in a real world problem is an open question. The error bars represent the standard deviation over 100 runs.}
\label{fig:B}
\end{figure}

>>>>

\usepackage{amssymb}
\usepackage{natbib}
\usepackage{graphicx}

@@

\title{Batch, Off-policy and Model-free Apprenticeship Learning}
%\author{Robert}
%   1. Supelec-Metz Campus, IMS Reasearch group, France. \texttt{prenom.nom@supelec.fr}}
\author{Edouard Klein\inst{1}\inst{3} \and Matthieu Geist\inst{1} \and Olivier Pietquin\inst{1}\inst{2}}
 \institute{
 1. Supélec-Metz Campus, IMS Research group, France, \texttt{prenom.nom@supelec.fr}\\
 2. UMI 2958 CNRS - GeorgiaTech, France\\
 3. Equipe ABC, LORIA-CNRS, France.
 }
\date{\today}

@@

\section{Background}
\label{sec:back}
A sequential decision problem is often framed as a Markov Decision
Process (MDP)~\citep{sutton1998reinforcement}, a tuple
$\{S,A,P,R,\gamma\}$ with $S$ being the state space, $A$ the action
space, $P\in\mathcal{P}(S)^{S\times A}$ the set of Markovian
transition probabilities, $R\in\mathbb{R}^S$ the reward function
(assumed to be absolutely bounded by 1) and $\gamma\in[0,1[$ a
discounting factor. A policy $\pi\in A^S$ maps states to actions. The
quality of a policy is quantified by the associated value function
$V^\pi$, which associates to each state the expected and discounted
cumulative reward: $V^\pi(s) = E[\sum\limits_{t=0}^\infty \gamma^t R(s_t)|s_0=s, \pi]$.
Dynamic programming aims at finding the optimal policy $\pi^*$, that
is one of the policies associated with the optimal value function,
$V^* = \argmax_\pi V^\pi$, which maximizes the value for each state.
If the model (that is transition probabilities and the reward
function) is unknown, learning the optimal control through
interactions is addressed by RL.

@@

We assume that the true reward function belongs to some hypothesis
space $\mathcal{H}_\phi = \{\theta^T \phi(s),
\theta\in\mathbb{R}^p\}$, of which we assume the basis functions to
be bounded by 1: $|\phi_i(s)|\leq 1, \forall s\in S, 1\leq i \leq
p$. Therefore, there exists a parameter vector $\theta^*$ such that:
 $R^*(s) = (\theta^*)^T \phi(s)$
In order to ensure that rewards are bounded, we impose that
$\|\theta\|_2\leq 1$. For any reward function belonging to
$\mathcal{H}_\phi$ and for any policy $\pi$, the related value
function $V^\pi(s)$ can be expressed as follows:
\begin{equation}
\scriptstyle
V^\pi(s) = E[\sum\limits_{t=0}^\infty \gamma^t \theta^T \phi(s_t)|s_0=s, \pi] = \theta^T  E[\sum\limits_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{equation}
Therefore, the value function is also linearly parameterized, with
the same weights and with basis functions being grouped into the
so-called \emph{feature expectation} $\mu^\pi$:
\begin{equation}
  \mu^\pi(s) = E[\sum_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{equation}
Recall that the problem is to find a policy whose performance is
close to that of the expert's for the starting state $s_0$, on the
unknown reward function $R^*$. In order to achieve this goal, it is
proposed in~\citep{abbeel2004apprenticeship} to find a policy $\tilde{\pi}$ such
that $\|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2\leq \epsilon$
for some (small) $\epsilon>0$. Actually, this ensures that the value
of the expert's policy and the value of the estimated policy (for
the starting state $s_0$) are close for \emph{any} reward function
of $\mathcal{H}_\phi$:
\begin{equation}
\scriptstyle
  |V^{\pi_E}(s_0) - V^{\tilde{\pi}}(s_0)| =
  |\theta^T(\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0))|
  \leq \|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2
\label{eqn:vs0}
\end{equation}
This equation uses the Cauchy-Schwarz inequality and the
assumption that $\|\theta\|_2\leq 1$. Therefore, the approach
described here does not ensure to retrieve the true reward function
$R^*$, but to act as well as the expert under this reward function
(and actually under any reward function).

@@

As suggested in~\citep{abbeel2004apprenticeship}, solving the MDP can be done
approximately by using any appropriate reinforcement learning
algorithm. In this paper, we use the Least-Squares Policy Iteration
(LSPI) algorithm~\citep{lagoudakis2003least}. There remains to estimate
feature expectations. In~\citep{abbeel2004apprenticeship}, $\mu^{\pi_E}(s_0)$ is
estimated using a Monte Carlo approach over $m$ trajectories:
$  \hat{\mu}_E(s_0) = \frac{1}{m} \sum\limits_{h=1}^m \sum\limits_{t=0}^\infty
  \gamma^t \phi(s_t^{(h)})$
This approach does not hold if only transitions of the expert are
available, or if trajectories are too long (in this case, it is
still possible to truncate them). For intermediate policies, it is
also suggested to estimate associated feature expectations using a
Monte Carlo approach (if they cannot be computed exactly). This is
more constraining than for the expert, as this assumes that a
simulator of the system is available. In order to address these
problems, we introduce a temporal-difference-based algorithm to
estimate these feature expectations.

\section{LSTD-$\mu$}
\label{sec:lstdmu}
Let us write the definition of the $i^\text{th}$ component of a
feature expectation $\mu^\pi(s)$ for some policy $\pi$:
$  \mu_i^\pi(s) = E[\sum\limits_{t=0}^\infty \gamma^t \phi_i(s_t)|s_0=s,\pi]$
This is exactly the definition of the value function of the policy
$\pi$ for the MDP considered with the $i^\text{th}$ basis function
$\phi_i(s)$ as the reward function. There exist many algorithms to
estimate a value function, any of them can be used to estimate
$\mu_i^\pi$. Based on this remark, we propose to use specifically
the least-squares temporal difference (LSTD)
algorithm~\citep{bradtke1996linear} to estimate each component of the
feature expectation (as each of these components can be understood
as a value function related to a specific and known reward
function).

Assume that a set of transitions $\{(s_t,r_t,s_{t+1})_{1\leq t \leq
n}\}$ sampled according to the policy $\pi$ is available. We assume
that value functions are searched for in some  hypothesis space
$\mathcal{H}_\psi = \{ \hat{V}_\xi(s) = \sum\limits_{i=1}^q \xi_i \psi_i(s)
= \xi^T \psi(s), \xi\in\mathbb{R}^q\}$. As reward and value
functions are possibly quite different, another hypothesis space is
considered for value function estimation. But if $\mathcal{H}_\phi$
is rich enough, one can still consider
$\mathcal{H}_\psi=\mathcal{H}_\phi$. Therefore, we are looking for
an approximation of the following form:
$  \hat{\mu}_i^\pi(s) = (\xi_i^*)^T \psi(s)$
The parameter vector $\xi_i^*$ is here the LSTD estimate:
\begin{equation}
  \xi_i^* = \left(\sum_{t=1}^n
  \psi(s_t)(\psi(s_t)-\gamma\psi(s'_{t}))^T\right)^{-1}
  \sum_{t=1}^n \psi(s_t) \phi_i(s_t)
\end{equation}
For apprenticeship learning, we are interested more particularly in
$\hat{\mu}^\pi(s_0)$. Let $\Psi = (\psi_i(s_t))_{t,i}$ be the
$n\times q$ matrix of values predictors, $\Delta\Psi = (\psi_i(s_t)
- \gamma\psi_i(s'_t))_{t,i}$ be the related $n\times q$ matrix and
$\Phi = (\phi_i(s_t))_{t,i}$ the $n\times p$ matrix of reward
predictors. It can be easily checked that $\hat{\mu}^\pi(s_0)$
satisfies:
\begin{equation}
  (\hat{\mu}^\pi(s_0))^T = \psi(s_0)^T (\Psi^T \Delta\Psi)^{-1} \Psi^T \Phi
\end{equation}
This provides an efficient way to estimate the feature expectation
of the expert in $s_0$.

There remains to compute the feature expectations of intermediate
policies, which should be done in an off-policy manner (that is
without explicitly sampling trajectories according to the policy of
interest). To do so, still interpreting each component of the
feature expectation as a value function, we introduce a state-action
feature expectation defined as follows (much as the classic
$Q$-function extends the value function):
$\mu^\pi(s,a) = E[\sum\limits_{t=0}^\infty \gamma^t
  \phi(s_t)|s_0=s,a_0=a,\pi]$
Compared to the classic feature expectation, this definition adds a
degree of freedom on the first action to be chosen before following
the policy $\pi$. With a slightly different definition of the
related hypothesis space, each component of this feature expectation
can still be estimated using the LSTD algorithm (namely using the
LSTD-Q algorithm~\citep{lagoudakis2003least}). The clear advantage of
introducing the state-action feature expectation is that this
additional degree of freedom allows off-policy learning.
Extending LSTD-$\mu$ to state-action LSTD-$\mu$ is done in the same
manner as LSTD is extended to LSTD-Q~\citep{lagoudakis2003least},
technical details are not provided here for the clarity of
exposition.

@@

We want to compare the efficiency of both versions of the algorithm with respect to the number of samples available from the expert, as these samples usually are the bottleneck. Indeed as they are quite costly (in both means and human time) to generate they are often not in abundance hence the critical need for an algorithm to be expert-sample efficient. Having a simulator at one's disposal can also be difficult. For the simple problem we use this is however not an issue. The discussion about the choice of the performance metric has its own dedicated subsection (subsection~\ref{ssec:quality}). We use here the $||\mu^{\pi^{(j)}}(s_0) - \mu^{\pi_E}(s_0)||_2$ error term. Fig.~\ref{fig:E} shows, for some numbers of samples from the expert, the value of $||\mu^{\pi^{(j)}}(s_0)-\mu^{\pi_E}(s_0)||_2$ after the algorithm converged or attained the maximum number of iterations (we used 40). The best policy is found by LSTD variant after one iteration only\footnote{Precise reasons why it happens are not clear now, but certainly have something to do with the fact that all the estimations are made along the same distribution of samples.} whereas in the MC variant, convergence happens after at least ten iterations. The best policy is not always the last, and although it has always been observed so far with the LSTD variant, there is absolutely no way to tell whether this is a feature. The score of the best policy (not the last) according to the $||\mu^{\pi^{(j)}}(s_0) - \mu^{\pi_E}(s_0)||_2$ error term is plotted here. We can see that although the LSTD variant is not as good as the MC variant when very few samples are available, both algorithms quickly converge to almost the same value ; our version converges to a slightly lower error value. The fact that our variant can work in a batch, off-policy and model-free setting should make it suitable to a range of tasks inaccessible to the MC variant, the requirement of a simulator being often constraining.
\subsubsection{Inverted Pendulum}
\begin{figure}
\begin{minipage}[t]{.4\linewidth}
    \begin{center}
      \resizebox{\columnwidth}{!}{\input{../Code/InvertedPendulum/threshold}}
      \caption{Number of balancing steps for the policies found by both variants with respect to the number of samples from a sub-optimal policy given to LSPI. Note the absence of middle ground in the quality.}
      \label{fig:threshold}
    \end{center}
\end{minipage}
\hfill
\begin{minipage}[t]{.4\linewidth}
    \begin{center}
      \resizebox{\columnwidth}{!}{\input{../Code/InvertedPendulum/threshold_EB}}
      \caption{Same plot as Fig.~\ref{fig:threshold}, with mean and standard deviation over 100 runs. The big standard deviation stems from the absence of middle ground in the quality of the policies. The increasing mean w.r.t. the abscissa means that the proportion of good policies increases with the number of samples from a sub-optimal policy given to LSPI.}
      \label{fig:threshold_EB}
    \end{center}
\end{minipage}
\end{figure}

%% \begin{multicols}{2}
%% \begin{figure}
%% \centering
%% \end{figure}
%% \begin{figure}
%% \centering
%% \end{figure}
%% \end{multicols}
Another classic toy problem is the \emph{inverted pendulum}, used for example in \citep{lagoudakis2003least} from where we drew the exact configuration (mass of the pendulum, length of the arm, definition of the feature space, etc.). In this setting, the machine must imitate an expert who maintains the unstable equilibrium of an inverted pendulum allowed to move along only one axis. A balancing step is considered successful (reward $0$) when the angle is less than $\pi\over 2$. It is considered a failure (reward $-1$) when the pendulum has fallen (i.e. the angle exceeds $\pi \over 2$). A run is stopped after a failure or after 3000 successful balancing steps.\\

@@

Fig.~\ref{fig:threshold_EB} shows that when given samples from a sub-optimal policy, the LSTD variant can sort the problem out, statistically almost always with 3000 sub-optimal samples, and sometimes with as low as 200 sub-optimal samples. Such a setting is still batch, off-policy and model-free. When given enough sub-optimal samples, both variants are successful with only one trajectory from the expert (i.e. 100 samples, this is what is plotted here). Giving more trajectories from the expert does not improve the success rate.
\subsection{Discussion about the quality criterion}
\label{ssec:quality}
\begin{figure}
\begin{minipage}[t]{.4\linewidth}
    \begin{center}
      \resizebox{\columnwidth}{!}{\input{../Code/GridWorld/criteria_mc}}
      \caption{Different criteria with respect to the number of iterations for a run of the MC variant.}
      \label{fig:A}
    \end{center}
\end{minipage}
\hfill
\begin{minipage}[t]{.4\linewidth}
    \begin{center}
\resizebox{\columnwidth}{!}{\input{../Code/GridWorld/criteria_lstd_EB}}
\caption{Different criteria with respect to the number of samples from the expert, for several runs of the LSTD variant. We can see that the algorithm is blind, as all it has access to is always zero. The true error values, however, smoothly converge to something small. Knowing how many expert samples are actually needed in a real world problem is an open question. The error bars represent the standard deviation over 100 runs.}
\label{fig:B}
    \end{center}
\end{minipage}
\end{figure}
